### 2018-05-04 papers/synth/example.tex
somya added
paper update

  %pm <-> ee
  \path (ee) edge[bend left=10] node [above] {\^{R3}} (pm);
  \path (pm) edge[bend left=30] node [above,rotate=-45] {\^{R2} {Qa7} Qbc7, R7, Qc7, Qa2} (ee);

@@

In figure~\ref{fig:mukund-vts}

The figure~\ref{fig:mukund} represent mammalian SNARE map created by studying the references []. In this diagram the gray circles represent compartments, the identities of compartments are
written within: ER=endoplasmic reticulum, ERGIC=ER-Golgi intermediate compartment,
RE=recycling endosome, EE=early endosome, LE=late endosome, LYS=lysosome, PM=plasma
membrane. Thin black arrows represent vesicle edges.
% and hollow arrows represent maturation edges. 
The set of SNAREs contained in these edges are written alongside each edge. Actual names
of these SNAREs are mentioned in the key alongside. The red labels represent the paths taken by
SNARES of the complex Stx-4-SNAP23-VAMP7. Here, the cycles for two SNARE molecules,
VAMP7 and SNAP23 are complete. The green labels represent the paths taken by SNAREs of the
complex Stx13-SNAP25-VAMP2. Here the cycle for VAMP2 is complete.The blue labels represent
the paths taken by SNARES of the complex Stx5-Gs28-Bet1/GS15-Ykt6. Here, cycles for none of
the SNAREs are complete.

>>>>

  %pm <-> ee
  \path (ee) edge[bend left=10] node [above] {\^{R3}} (pm);
  \path (pm) edge[bend left=30] node [above,rotate=-45,text width = 2.5cm] {\^{R2} {Qa7} Qbc7, R7, Qc7, Qa2} (ee);

@@

%t snares are labelled within compartmetns, v snares on the edges
%and snare complexes on the left-hand side
The figure~\ref{fig:mukund-vts} represent mammalian SNARE map created by studying the references~\cite{somya}. 
To construct the map, we have assumed that vesicles only contain a single
active v-SNARE, and we have attributed t-SNAREs and inactive v-SNAREs that
travel between compartments to one of the known vesicles that go between
the same source and target compartments.
In order to identify the active SNARE complex involved in any particular
vesicle fusion, two criteria were used: 
\begin{enumerate}
\item[a.] The SNARE complex is formed \textit{in vivo}. In most papers, this is determined by immunoprecipitation of the SNARE complex from the relevant cell fraction. 
\item[b.] Blocking SNARE complex formation (for example, using antibodies against these SNAREs, or
using cytosolic forms of these SNAREs) blocks the specific transport step.
\end{enumerate}

Note that these vesicles have been collected from multiple cell types, and
any given cell type is likely to contain only a subset of the vesicles in
the map.

In this diagram the gray circles represent compartments, the identities of compartments are
written within: ER=endoplasmic reticulum, ERGIC=ER-Golgi intermediate compartment,
RE=recycling endosome, EE=early endosome, LE=late endosome, LYS=lysosome, PM=plasma
membrane. Thin black arrows represent vesicle edges.
% and hollow arrows represent maturation edges. 
The set of SNAREs contained in these edges are written alongside each edge. Actual names
of these SNAREs are mentioned in the key alongside. The red labels represent the paths taken by
SNARES of the complex Stx-4-SNAP23-VAMP7. Here, the cycles for two SNARE molecules,
VAMP7 and SNAP23 are complete. The green labels represent the paths taken by SNAREs of the
complex Stx13-SNAP25-VAMP2. Here the cycle for VAMP2 is complete.The blue labels represent
the paths taken by SNARES of the complex Stx5-Gs28-Bet1/GS15-Ykt6. Here, cycles for none of
the SNAREs are complete. So, even though most edges known so far are present in the map, paths for molecules (SNAREs) across the cell are still not complete.

We have used our tool to synthesize possible fixes to complete the graph. Variation 1 of our tool adds new labelled edges to the graph. We have ranked the all possible graph edits with simple rank of minimum updates. Our tools was able to complete this graph by adding five new edges and nine new molecules in 210s. Variation 2 of our graph adds new labels to the edge and was able to fix this graph with eleven new molecule labels in 180s. Variation 3 of our tool synthesis low depth CNF functions and was able to synthesis the result in 500s. Variation 4 of our tool synthesis Boolean function based on AND, OR gates. The result shows the basic template of 4 leafs and 3 gates we allow user to provide it's own template in our front end language. The tool was able to synthesize the Boolean functions in 470s.

### 2018-05-04 papers/synth/experiments.tex
paper update

We implemented the encodings in a tool called~\ourtool\footnote{\ashu{\url{http://add-git-hub-url}}}.
%
The tool takes a partially defined VTS as input in its custom designed
input language.
%
The tool can not only synthesize the above discussed query, but also their
combination.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
Our tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure for
processing formulas. Since the formulas are in QBF, we cannot use~\zthree.
%
We use~\depqbf~\cite{depqbf} for solving of QBF formulas.
%
Our tool includes about ?? lines of code.

We have applied~\ourtool~on six partially defined VTS.
%
The results are presented in table~\ref{tab:qf-graph}.
%
The experiments were conducted on ... machine,
with \ashu{?}GB memory and 900s timeout.
%
The first four VTSs are synthetic and inspire from literature for typical
motifs in VTSs. 
%
The last two are the VTSs that are from .......\ashu{Describe the source of two examples}.

>>>>

We implemented the encodings in a tool called~\ourtool\footnote{{\url{https://github.com/arey0pushpa/pyZ3}}}.
%
The tool takes a partially defined VTS as input in its custom designed
input language. The input is then converted to the constraints over VTS. 
%
The tool can not only synthesize the above discussed query, but also their
combination.
%
For example, our tool can modify labels of nodes or edges while
learning activity functions.
%
Our tool is developed in C++ and uses~\zthree~\cite{z3} infrastructure for
processing formulas. 
%
Since, the formulas involve alternation of quantifiers over Boolean variables Z3 is not a suitable choice. We translate the formula created by Z3 tool into standard QDIMACS~\cite{qdimacs} format. Which can be used as an input for most of the QBF solvers.

%Since the formulas are in QBF, we cannot use~\zthree.
%
We use~\depqbf~\cite{lonsing2010depqbf} for solving of QBF formulas. 
%
Our tool includes about 7000 lines of code.

We have applied~\ourtool~on six partially defined VTS.
%
The results are presented in table~\ref{tab:qf-graph}.
%
The experiments were conducted on ... machine,
with \ashu{?}GB memory and 900s timeout.
%
The first four VTSs are synthetic and inspire from literature for typical
motifs in VTSs. 
%
%
The fifth VTS is taken from "A complete set of SNAREs in yeast"
~\cite{burri2004complete} paper. It has been adapted from the paper by separating the v and the t SNAREs. 
%The t SNAREs are labelled within compartments and v SNAREs on the edges. %\ashu{Describe the source of two examples}.
%
The last VTS represent mammalian SNARE map created by studying the literature references.

### 2018-05-04 papers/synth/intro.tex
paper update

%Motivate VTS
\ashu{Reduce the intro by half page}

Eukaryotic cells, including human cells, consist of multiple membrane
bound compartments.
%
Material is transported among these compartments by the vesicle
transport system (VTS) [].
%
Briefly, the source compartment produces a membrane bound packet of
molecules called a vesicle.
%
After release, this vesicle specifically recognizes the correct target
compartment within the cell, and fuses with it.
%
A lot of information about the molecules that form the machinery of
the VTS has been discovered, including their regulatory interaction
with each other. 
% %
% \todo{connect to function }{We also know a lot about the specific functions} of these molecules and how these molecules interact amongst each other.
%
In spite of this detailed knowledge at the level of the molecules, the
structure of the VTS network, or the road-map of the eukaryotic cell,
is far from complete.
%
The current knowledge of the network is put together from a patch work
of biological experiments and is scattered across several publications.
%
Even after this information is collected and put together, we find
that the network obtained is still not complete; new vesicles and new
contents in previously known vesicles are constantly being discovered.
%
The synthesis for the unknown pieces may be assisted by computation on
the graph model of VTSs.
%
% 
%In this work, we use a model described in [] to find out, given a partial
%VTS graph and details about interactions amongst its molecules, what the complete graph of the
%VTS could look like. (Please look at appendix A for the detailed discussion on pros and cons of the
%model.
%
%Eukaryotic cells, including human cells, consists of multiple compartments.
%
%Vesicle Traffic Systems(VTSs) are the material transport mechanisms among the compartments inside the cells~\cite{vtsIntro}.
%
%Almost all subsystems of the cells depend on VTSs.
%
%Therefore, understanding how VTSs functions is one of the key
%questions of cell biology.
%
In this paper, we are looking at the computational questions 
arise from the VTSs.

We use the model of VTSs that has been presented in~\cite{VTS}.
%
Please look at appendix~\ref{sec:model}
for the detailed discussion on pros and cons of the model.
% Describe VTS in detail
While VTSs transports molecules, VTSs are also
regulated by the molecules.
%
Each compartment contains a set of molecules.
%
The molecules are transported to the other compartments via
unidirectional channels.  
%
The molecules in a compartment may be active or not active.
%
Similarly, a molecule can be active or inactive in a channel. 
%
The active molecules are the regulators of the channels.
%
A channel is enabled by a pair of sets of molecules
such that one occurs in the channel and the other
occurs in the destination compartment.
%
The molecules in both the sets must be active in the channel
and compartment respectively.
%
The pairs are called {\em fusing} sets and analogously
the channel is considered to be {\em fused} with the
destination compartment.
%
Not all the pairs of set of molecule can participate in the fusion, the
potential fusion pairs are usually determined by biological experiments.
%
To ensure effective regulation, the fusing pairs must be
unique for each fused channel-compartment pair.
%
A molecule that has participated in a fusion must not interfere
with fusion at different compartments or channels.
%
Therefore, the molecule must be inactive in appropriate compartments.
%
The activity of molecules is regulated by the other molecules, i.e.,
the presence and absence of the other molecules in a compartment or
channel may make the molecule active or inactive.
%
We call this regulation as {\em activity functions}.

% Defending our model
%
In the VTS model, we assume that the system is in study state and the
concentrations of the molecules in the compartments do not change over
time.
%
We model the system as a labelled graph, where compartments are nodes and
transport channels are edges.
%
The set of molecules present and their activity in a compartment or
a channel is the label of the respective node or edge.
%
The regulation controls are defined by a fusion pairing relation
containing pairs of molecules and activity
Boolean functions.
%
Since the system is in study state, we expect that any molecule that
leaves a compartment must come back via some path on the graph.
%
We call this property of VTS as {\em stability}.

% Partial info defence
%
As we have discussed earlier, our understanding of VTSs are partial.
%
% It largely means that the current knowledge is due to the patch work
% of biological experiments~\cite{model} for identifying the compartments,
% channels, molecules, and their activity and regulatory control.
% %
% Often the information is scattered around in several publications.
% %
% Even after with some effort we may collate all the known information about
% a VTS, we may still have unknown pieces of the system; new vesicles and new contents in previously known vesicles are constantly being discovered.
%
%{Speculative: Make a case for completing partial parts of VTSs!!}
%
For example, we may suspect that some edges may not have been
discovered in a VTS that does not satisfy the stability condition.
%
The synthesis for the unknown pieces may be {\em assisted} by computation on
the graph model of VTSs.
%
In this paper, we consider several versions of the synthesis problem
involving different parts of VTSs that can be synthesized, such as
modifying labels, adding/deleting edges, and learning activity functions.
%
We also consider variation on the properties against which we do
synthesis, namely stability, and $k$-connectedness that states that
VTS remains connected after removing any $k-1$ edges.
%
In order to synthesize the parts of a VTS such that it satisfies the
constraints, we encode the synthesis problem into a satisfiability of
quantified Boolean formulas(QBFs). 
%

We have implemented the encoding in a flexible tool,
which can handle a wide range of synthesis query.
%
We have applied our tool on two found-in-nature VTSs, namely
M1~\cite{} and S1~\cite{}.
%
We have also applied our tool on various synthetic examples to
demonstrate that our tool scale upto the graph size ??, which
is the size of typical VTSs in a eukaryotic cell.

>>>>

%Motivate VTS
\ashu{reduce the intro by half page}

Eukaryotic cells, including human cells, consist of multiple membrane
bound compartments.
%
Material is transported among these compartments by the vesicle
transport system (VTS).
%
Briefly, the source compartment produces a membrane bound packet of
molecules called a vesicle.
%
After release, this vesicle specifically recognizes the correct target
compartment within the cell, and fuses with it.
%
A lot of information about the molecules that form the machinery of
the VTS has been discovered, including their regulatory interaction
with each other. 
% %
% \todo{connect to function }{We also know a lot about the specific functions} of these molecules and how these molecules interact amongst each other.
%
In spite of this detailed knowledge at the level of the molecules, the
structure of the VTS network, or the road-map of the eukaryotic cell,
is far from complete.
%
The current knowledge of the network is put together from a patch work
of biological experiments and is scattered across several publications.
%
Even after this information is collected and put together, we find
that the network obtained is still not complete; new vesicles and new
contents in previously known vesicles are constantly being discovered.
%
The synthesis for the unknown pieces may be assisted by computation on
the graph model of VTSs.
%
% 
%In this work, we use a model described in [] to find out, given a partial
%VTS graph and details about interactions amongst its molecules, what the complete graph of the
%VTS could look like. (Please look at appendix A for the detailed discussion on pros and cons of the
%model.
%
%Eukaryotic cells, including human cells, consists of multiple compartments.
%
%Vesicle Traffic Systems(VTSs) are the material transport mechanisms among the compartments inside the cells~\cite{vtsIntro}.
%
%Almost all subsystems of the cells depend on VTSs.
%
%Therefore, understanding how VTSs functions is one of the key
%questions of cell biology.
%
In this paper, we are looking at the computational questions 
arise from the VTSs.

We use the model of VTSs that has been presented in~\cite{VTS}.
%
Please look at appendix~\ref{sec:model}
for the detailed discussion on pros and cons of the model.
% Describe VTS in detail

VTSs are regulated by the same molecules that they transport. Molecules on compartments and on vesicles can be in either an active or an inactive state.
% the term vesicle has already been introduced.. using two terms for the same object seems confusing. 
%
% (The molecules are transported to the other compartments via
%unidirectional channels.) cannot comment on unidirectionality of vesicles. they may be undergoingmultiple cycles of back-fusion for all we %know..  
%
The active molecules regulate the vesicle they occur on.
%
A vesicle is enabled by a set of molecules
such that one part of the set occurs in the vesicle and the other part
occurs in the target compartment.
% use either target or destination consistantly throughout..i prefer target because that's what the t in t-SNARE stands for.
%
These molecules must be active in the vesicle
and target compartment respectively.
%
The pairs are called {\em fusing} sets and analogously
the vesicle is considered to be {\em fused} with the
destination compartment.
%
Not all sets molecule can participate in the fusion, the
potential fusion pairs are usually determined by biological experiments.
%
Generally, in cells, the fusing pairs are found to be distinct for distinct vesicle-compartment fusions.
%
In order that a molecule that has participated in a fusion does not interfere
with fusion at compartments and vesicles, in the model, we require that the molecule is inactive on appropriate compartments.
%
%
The activity of molecules is regulated by the other molecules, i.e.,
the presence and absence of the other molecules in a compartment or
vesicle may make the molecule active or inactive.
%
We call this regulation as {\em activity functions}.

% Defending our model
%
In this VTS model, we assume that the system is in steady state and the
concentrations of the molecules in the compartments do not change over
time.
%
We model the system as a labelled graph, where compartments are nodes and
transport vesicles are edges.
%
The set of molecules present and their activity in a compartment or
a vesicle is the label of the respective node or edge.
%
The regulation controls are defined by a fusion pairing relation
containing pairs of molecules and activity
Boolean functions.
%
Since our system is in steady state, we expect that any molecule that
leaves a compartment must come back via some path on the graph.
%
We call this property of VTS as {\em stability}.

% Partial info defence
%
As we have discussed earlier, our understanding of VTSs is partial.
%
% It largely means that the current knowledge is due to the patch work
% of biological experiments~\cite{model} for identifying the compartments,
% channels, molecules, and their activity and regulatory control.
% %
% Often the information is scattered around in several publications.
% %
% Even after with some effort we may collate all the known information about
% a VTS, we may still have unknown pieces of the system; new vesicles and new contents in previously known vesicles are constantly being discovered.
%
%{Speculative: Make a case for completing partial parts of VTSs!!}
%
%For example, it is possible that some edges may not have been
%discovered in a VTS that does not satisfy the stability condition.
% nature of missing information has already been elaborated upon in the first paragraph
%
The synthesis for the unknown pieces may be {\em assisted} by computation on
the graph model of VTSs.
%
In this paper, we consider several versions of the synthesis problem
involving different parts of VTSs that can be synthesized, such as
modifying labels, adding/deleting edges, and learning activity functions.
%
We also consider variation on the properties against which we do
synthesis, namely stability, and $k$-connectedness that states that
VTS remains connected after removing any $k-1$ edges.
%
In order to synthesize the parts of a VTS such that it satisfies the
constraints, we encode the synthesis problem into a satisfiability of
quantified Boolean formulas(QBFs). 
%

We have implemented the encoding in a flexible tool,
which can handle a wide range of synthesis query.
%
We have applied our tool on two found-in-nature VTSs, namely
M1~\cite{} and Y1~\cite{burri2004complete}.
%
We have also applied our tool on various synthetic examples to
demonstrate that our tool scale upto the graph size ??, which
is the size of typical VTSs in a eukaryotic cell.

### 2018-05-04 papers/synth/model.tex
paper update

The molecules transported by the VTS are themselves its regulators. The molecules in a
compartment/vesicle may be active or inactive. The molecules that are respnsible for vesicle fusion
are called SNARE proteins []. Active SNAREs present on vesicles (v-SNAREs) bind with their
cognate active SNAREs on the target compartment (t-SNAREs) to enable vesicle fusion. A cell
contains multiple kinds of v- and t-SNAREs. Only specific pairs of v and t SNAREs can bind to
each other and participate in fusion. Fusion compatible v- and t- SNAREs are determined by
biological experiments. Different vesicle-compartment fusions in the cell are brought about by
different v- and t-SNARE pairs. A molecule that participates in a given fusion reaction must not
interfere with fusion at different compartments or vesicles. Therefore, SNAREs must be kept in an
inactive form in appropriate compartments/vesicles. The activity of molecules is regulated by the
other molecules, i.e., the presence and absence of the other molecules in a compartment or vesicle
may make the molecule active or inactive. We call this regulation as activity functions.
In the VTS model, we assume that the system is in steady state and the concentrations of the
molecules in the compartments do not change over time. We define SNARE pairing specificity by a
fusion pairing relation containing pairs of SNAREs and molecular regulation by activity Boolean
functions. Since the system is in steady state, we expect that any molecule that leaves a
compartment must come back via some path on the graph. We call this property of VTS as stability.

A few assumptions on modeling level.
\begin{enumerate}
\item[1] A cell is a set of compartments exchanging vesicles.
\item[2] Compartments are neither created nor destroyed.
\item[3] Each compartment is in steady state, gain and loss balance.
\item[4] Molecules are neither created nor destroyed.
\item[5] Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}

\ankit{Discuss the yeast and eukaryotes VTS model and assumptions }

\begin{enumerate}
\item Complex form using three Q-SNAREs and one R-SNARE. In order for a vesicle to fuse to it's corresponding target compartment this combination has to be present.
\item In the graph shown the R-SNAREs are always in active. Every combination is formed by three active Q-SNAREs and one active R-SNARE.
\item  Qai, Qbi, Qci and Ri are just dummy names, real biological names can be found in the below table. 
\item In the diagram white circle molecules are always active. Other edge labels are inactive molecules.
\item Pairing matrix restriction: ith R-SNARE can fuse with only the corresponding ith Q-SNARES and vice-versa. No other combination is allowed.
\end{enumerate}

@@

Qa1 Qb1 Qc1 R1
Stx18 Sec20 Slt1 Sec22b

>>>>

The molecules transported by the VTS are themselves its regulators. The molecules in a
compartment/vesicle may be active or inactive. The molecules that are responsible for vesicle fusion
are called SNARE proteins []. Active SNAREs present on vesicles (v-SNAREs) bind with their
cognate active SNAREs on the target compartment (t-SNAREs) to enable vesicle fusion. A cell
contains multiple kinds of v- and t-SNAREs. Only specific pairs of v and t SNAREs can bind to
each other and participate in fusion. Fusion compatible v- and t- SNAREs are determined by
biological experiments. Different vesicle-compartment fusions in the cell are brought about by
different v- and t-SNARE pairs. A molecule that participates in a given fusion reaction must not
interfere with fusion at different compartments or vesicles. Therefore, SNAREs must be kept in an
inactive form in appropriate compartments/vesicles. The activity of molecules is regulated by the
other molecules, i.e., the presence and absence of the other molecules in a compartment or vesicle
may make the molecule active or inactive. We call this regulation as activity functions.
In the VTS model, we assume that the system is in steady state and the concentrations of the
molecules in the compartments do not change over time. We define SNARE pairing specificity by a
fusion pairing relation containing pairs of SNAREs and molecular regulation by activity Boolean
functions. Since the system is in steady state, we expect that any molecule that leaves a
compartment must come back via some path on the graph. We call this property of VTS as stability.

Our model is inspired by~\cite{shukla2017discovering}. On the timescales of minutes, our following assumptions reasonably capture the important aspects of the Rothman-Schekman-Sudhof (RSS) model~\cite{rothman2002machinery} of vesicle traffic system.
\begin{enumerate}
\item[1] A cell is a set of compartments exchanging vesicles.
\item[2] Compartments are neither created nor destroyed.
\item[3] Each compartment is in steady state, gain and loss balance.
\item[4] Molecules are neither created nor destroyed.
\item[5] Molecules move via vesicles of uniform size.
\item Identical vesicles have identical target compartments.
\item Fusion of vesicles to compartments is driven by specific SNARE pairing.
\item The activity of a SNARE can be regulated by other molecules present on the same compartment
or vesicle.
\item An active SNARE pair is necessary and sufficient for fusion. 
\end{enumerate}

SNARE proteins are the agents of vesicle fusion in eukaryotic cells. When SNAREs on vesicles (v-SNAREs) encounter their cognate SNAREs on target compartments (t-SNAREs), they form SNARE complexes~\cite{jahn2006snares}, and a single SNARE complex releases enough energy to enable membrane fusion~\cite{van2010one}. SNAREs are identified by the presence of a conserved 60-70 stretch of amino acids called the SNARE motif. Based on amino acid sequence, SNARE motifs fall into 4 classes: Qa, Qb, Qc, and R~\cite{jahn2006snares}. Across all intracellular vesicle fusion reactions, the associated SNARE complexes contain one of each of the four kinds of SNARE motifs; the v-SNARE contributes a single SNARE motif, usually it is an R-SNARE (although, exceptions are known: Sec22b and Ykt6 are both R-SNAREs which form parts of t-SNAREs~\cite{hong}) and the rest of the three SNARE
motifs are contributed by the t-SNARE. In the cell, different vesicle fusion reactions are associated with distinct v- and t-SNARE pairs.

We have extended the model by considering each complex molecule as distinct. In the previous paper, we have considered three Q SNARES as a single molecule. We also do not assume that SNARE type distribution is even in the whole system. The fusion is now driven by the active presence of three Q SNARE and one R SNARE molecule, present in any combination on edge and on the target compartment. We have relaxed the pairing matrix constraint to comply with this fact. In order for the biological efficiency and optimality reasons we still do not allow self-edges to be present in the VTS. 
%\begin{enumerate} 
%\item Complex form using three Q-SNAREs and one R-SNARE. In order for a vesicle to fuse to it's corresponding target compartment this combination has to be present.
%\item In the graph shown the R-SNAREs are always in active. Every combination is formed by three active Q-SNAREs and one active R-SNARE.
%\item  Qai, Qbi, Qci and Ri are just dummy names, real biological names can be found in the below table. 
%\item In the diagram white circle molecules are always active. Other edge labels are inactive molecules.
%\item Pairing matrix restriction: ith R-SNARE can fuse with only the corresponding ith Q-SNARES and vice-versa. No other combination is allowed.
%\end{enumerate}

@@
